---
title: "Groceries Product Analysis with Apriori"
output: html_notebook
---

Import data set and read data set into data frame
```{r}
groceries <- read.csv("C:/Users/Jacque de l'est/Documents/Datasets for Data Science/Groceries_dataset.csv", encoding = "UTF-8",)
head(groceries, 10)
```

See descriptive statistics of data frame
```{r}
summary(groceries)
```

See missing values of data frame
```{r}
colSums(is.na(groceries))
```

Convert datatype of Member_number to numeric datatype and datatype of Date to date
```{r}
groceries$Member_number <- as.numeric(groceries$Member_number)
groceries$Date <- as.Date(groceries$Date, format = "%d-%M-%Y")
head(groceries)
```
```{r}
sorted_groceries <- groceries[order(groceries$Member_number),]
head(sorted_groceries, 10)
```

Group purchased items by customer ID and purchasing date
```{r}
library(plyr)
itemList <- ddply(sorted_groceries, c("Member_number","Date"), 
                  function(df1)
                      paste(df1$itemDescription, collapse = ","))
head(itemList, 10)
```

Delete columns Member_number and Date, rename column v1, and save to file
```{r}
clean_itemList <- drop(itemList[3])
colnames(clean_itemList) <- 'Items'

write.csv(clean_itemList, "ItemList.csv", quote=FALSE, row.names=TRUE)
head(clean_itemList, 10)
```

Reads a transaction data file and creates a transactions object
```{r}
library(arules)
groceries_tabular <- read.transactions("./ItemList.csv", format = "basket", cols = 1, sep = ",", skip = 1)
```
Show descriptive statistics of data frame
```{r}
summary(groceries_tabular)
```

Create Top 10 Groceries
```{r}
top10_groceries <- itemFrequency(groceries_tabular, type = "absolute")
top10_groceries <- sort(top10_groceries, decreasing = TRUE)[1:10]
top10_groceries <- data.frame("ProductName" = names(top10_groceries),
                              row.names = NULL)
head(top10_groceries, 10)
write.csv(top10_groceries, file="Top10_Groceries.txt")
```

Create Bottom 10 Groceries
```{r}
bottom10_groceries <- itemFrequency(groceries_tabular, type="absolute")
bottom10_groceries <- sort(bottom10_groceries, decreasing=FALSE)[1:10]
bottom10_groceries <- data.frame("ProductName" = names(bottom10_groceries),
                                 row.names=NULL)
head(bottom10_groceries)
write.csv(bottom10_groceries, file="bottom10_groceries.txt")
```

Create an item frequency plot for the top 10 items
```{r}
if (!require("RColorBrewer")) {
  #install color package of R
  install.packages("RColorBrewer")
  #include library RColorBrewer
  library(RColorBrewer)
}

itemFrequencyPlot(groceries_tabular, 
                  topN = 10, 
                  type = "absolute", 
                  col = brewer.pal(8,'Pastel2'), 
                  main = "Absolute Item Frequency Plot")
itemFrequencyPlot(groceries_tabular, 
                  topN = 10, 
                  type = "relative", 
                  col = brewer.pal(8,'Pastel2'), 
                  main = "Relative Item Frequency Plot")
```

Use apriori algorithm
```{r}
assorules <- apriori(groceries_tabular, 
                      parameter = list(support = 10/length(groceries_tabular), 
                                       minlen = 2, 
                                       conf = 0.05
                                       )
                      )
```

```{r}
summary(assorules)
```

```{r}
inspect(assorules[1:10])
```

Visualize the Association Rules
```{r}
library(arulesViz)
plot(assorules, jitter=0)
```

```{r}
plot(assorules, method="grouped", control=list(k=5))
```

```{r}
plot(assorules[1:20], method="graph")
```

```{r}
plot(assorules[1:20], method="paracoord")
```

